---
title: "DataFest R Workshop May 2020"
author: "Nathalie Moon (adapted from Sotirios Damouras' 2019 Workshop)"
date: "May 28, 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
```


# PART 1

## Task 1.1
Create an R project & associated folder for the workshop

## Task 1.2
Download and install the 'tidyverse' library

```{r, include=FALSE}
# install.packages("tidyverse") ## Note - you only need to do this once on your computer.  Once it is installed, you will only need to load it with the library() function


```

## Task 1.3
Download the "dinesafe.csv" file from the workshop folder and save it in a project subfolder called "data". Then read the data into a data-frame called "dinesafe" 

```{r}

```

## Task 1.4
Use the `glimpse()` function to take a look at the `dinesafe` data frame's structure
```{r}

```


## Task 1.5
Use the `View()` function to view the `dinesafe` data frame as a spreadsheet
```{r}

```


\newpage
# Part 2

## Task 2.1
Find all distinct establishment types. Hints: Are values of establishment in different rows or different columns?  Which of the `dplyr` functions can you use to remove duplicated values?



## Task 2.2
Find all inspections that took place on August 21st, 2018 (i.e. "2018-08-21").  Hints: (1) What variable contains the date of inspections? (2) Which `dplyr` function can you use to keep only observations for establishments that got inspected on this date?




  
## Task 2.3
Find the total # of distinct inspections 



## Task 2.4
Rank establishment types by total amount fined




## Task 2.5
Rank establishment types by average amount fined per establishment. Hints: (1) What variable might you want to group on?  (2) How can you calculate the average amount fined in each category?




## Task 2.6 (Challenging)
Find the establishment with the highest non-zero total fine amount whithin each establishment type.  Hints: (1) Start by calculating the total fine for each establishment (think of what variable to group on to achieve this), then create new groups based on `ESTABLISHMENTTYPE` and use one of the `dplyr` functions to keep only the observations with the highest total in each of these new groups.  Note - you'll need to use `ungroup` before regrouping the data into new groups.


\newpage
# PART 3
The file data/establishments.csv contains information on different establishments, in particular its neighborhood.


# We will try to match this information with the dinesafe data.
# Note that NOT ALL inspected establishments are present. 


## Task 3.1
Do an inner_join between the dinesafe and establishments tables. Hint: Which variable will you use to do the matching (it should be a variable which is present in both datasets)



## Task 3.2
Use inner_join to rank the neighborhoods by the number of "C - Crucial" type infractions in restaurants. Hint: Either before or after joining, you'll need to use the `filter` function to keep only the observations we're interested in here (e.g. restaurants with inspection results of "C - Crucial"). After joining, think about which variable to group by before sorting. 




  
## Task 3.3
Find which (distinct) establishments did NOT get matched to a neighborhood. Hint: Use `anti_join`




\newpage
# PART 4

Our goal is to create a publication-quality graph with ggplot2. You will try to reproduce the Gapminder World Poster on World Health: https://www.gapminder.org/downloads/updated-gapminder-world-poster-2015/ using data in the gapminder package

```{r}
# load data 
#install.packages("gapminder")
library(gapminder)

# Take a look at the data 
glimpse(gapminder)
```


## Task 4.1
Create a scatter-plot of Life Expectancy (lifeExp) versus GDP-per-capita (gdpPercap), for 2007 data




## Task 4.2
On your previous plot, change the x-axis (gdg/cap) to log-scale. Hint: look at the choices for "Scale" geometries on the `ggplot2` cheat sheet




## Task 4.3
On your previous plot, change change the size of the points according to the population of each country




## Task 4.4
On your previous plot, change the color of the points according to the continent




## Task 4.5
On your previous plot, change the scale of each point to range from 1 to 14. Hint: Use the `scale_size(range = )` geometry to specify the range of sizes




## Task 4.6
On your previous plot, label to each point with the name of the country; use the `geom_text` function with options `nudge_x = .02`, `alpha = .2` to make the labels readable - play around with values of `nudge_x` and `alpha` to figure out what these are doing!



## Task 4.7
Modify the previous plot to make facets for each continent.



## Task 4.8
If your x-axis labels are hard to read due to overlapping, you may want to rotate theme. You can use the `theme(axis.text.x=element_text(angle = 90, hjust = 0))` geometry to adjust this; change the values of `angle` and `hjust` to figure out the effect of these two arguments and find a combination that makes the labels easier to read.

